Shrinking Characteristics of Precision Matrix Estimators: An Illustration via Regression

### 2019-03-10 Manuscript.tex
Deploy MGallow/oral_manuscript to github.com/MGallow/oral_manuscript.git:gh-pages
added notations and restructured

\documentclass[12pt,]{book}

@@

@@

            pdftitle={Oral Manuscript},

@@

@@

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

@@

\usepackage{booktabs}
\usepackage{amsthm}
\makeatletter
\def\thm@space@setup{%
  \thm@preskip=8pt plus 2pt minus 4pt
  \thm@postskip=\thm@preskip
}
\makeatother
\usepackage[]{natbib}
\bibliographystyle{apalike}

\title{Oral Manuscript}
\author{Matt Galloway}
\date{2018-06-07}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{corollary}{Corollary}[chapter]
\newtheorem{proposition}{Proposition}[chapter]
\theoremstyle{definition}
\newtheorem{example}{Example}[chapter]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[chapter]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{prerequisites}{%
\chapter{Prerequisites}\label{prerequisites}}

\hypertarget{intro}{%
\chapter{Introduction}\label{intro}}

This is the introduction for Matt's oral manuscript. (UPDATED 8)

\hypertarget{tutorial}{%
\chapter{Tutorial}\label{tutorial}}

@@

>>>>

\documentclass[11pt,]{report}

@@

\usepackage{setspace}
\setstretch{1.5}

@@

            pdftitle={Shrinking Characteristics of Precision Matrix Estimators: An Illustration via Regression},

@@

\usepackage[left = 1in, right = 1in, top = 1in, bottom = 1in]{geometry}

@@

@@

\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{fancyhdr} % Required for custom headers
\usepackage{lastpage} % Required to determine the last page for the footer
\usepackage{extramarks} % Required for headers and footers
\makeatletter
\def\thm@space@setup{%
  \thm@preskip=8pt plus 2pt minus 4pt
  \thm@postskip=\thm@preskip
}
\makeatother
\pagestyle{fancy}
% \pagestyle{empty}
\usepackage[left]{lineno} % line number package
\linenumbers % line number command
\usepackage[]{natbib}
\bibliographystyle{apalike}

\title{Shrinking Characteristics of Precision Matrix Estimators: An Illustration via Regression}
\author{Matt Galloway}
\date{2019-03-09}

\begin{document}
\maketitle

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{1}
\tableofcontents
}
\hypertarget{intro}{%
\chapter{Introduction}\label{intro}}

In many statistical applications, estimating the covariance matrix, \(\Sigma\), for a set of random variables is a critical task. The covariance matrix is useful because it characterizes the \emph{relationship} between variables. For instance, suppose we have three variables \(X, Y, \mbox{ and } Z\) and their covariance matrix is of the form

\begin{equation}
\Sigma_{xyz} = \begin{pmatrix}
1 & 0 & 0.5 \\ 
0 & 1 & 0 \\ 
0.5 & 0 & 1
\end{pmatrix}
\label{eq:ex}\notag
\end{equation}

We can gather valuable information from this matrix. First of all, we know that each of the variables has an equal variance of 1. Second, we know that variables \(X\) and \(Y\) are likely independent because the covariance between the two is equal to 0. This implies that any information in \(X\) is useless in trying to gather information about \(Y\). Lastly, we know that variables \(X\) and \(Z\) are moderately, positively correlated because their covariance is 0.5.

Unfortunately, estimating \(\Sigma\) well is often computationally expensive and, in a few settings, extremely challenging. For this reason, emphasis in the literature and elsewhere has been placed on estimating the inverse of \(\Sigma\) which we like to denote as \(\Omega \equiv \Sigma^{-1}\).

\hypertarget{references}{%
\subsection{references}\label{references}}

The reference we need here is \citet{molstad2017shrinking}!

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  From \citet{molstad2017shrinking}: ``To fit many predictive models, only a characteristic of the population precision matrix needs to be estimated. For example, in binary linear discriminant analysis, the population precision matrix is needed for prediction only through the product of the precision matrix and the difference between the two conditional distribution mean vectors. Many authors have proposed methods that directly estimate this characteristic \citet{cai2011direct}; \citet{fan2012road}; \citet{mai2012direct}.''
\item
  From \citet{molstad2017shrinking}: ``\citet{chen2016regularized} proposed a method for estimating the characteristic \(\Omega_{*}\mu_{*}\) directly, but like the direct linear discriminant methods, this apporach does not lead to an estimate of \(\Omega_{*}\).''
\item
  Discuss linear regression application\ldots{}. where \(\beta = \Omega_{*}\Sigma_{*XY}\). Like \citet{witten2009covariance}, this approach estimates the regression coefficients ``using shrinkage estimators of the marginal precision matrix for the predictors''.
\end{enumerate}

\begin{itemize}
\item
  Guide
\item
  Model and notation
\end{itemize}

\hypertarget{outline}{%
\section{Outline}\label{outline}}

\begin{itemize}
\item
  Why do we care about covariance matrices? Covariance matrices are an object that summarizes the relationship between variables.
\item
  What is a precision matrix? Precision matrix is the inverse of a covariance matrix.
\item
  Precision matrices are powerful, essential in representing relationships in variables
\item
  In fact, required to be estimated to fit many statistical models.
\item
  That is, we must provide an object that accurately summarizes the relationship and interaction between variables in our model.
\item
  Many proposed shrinkage estimators when p is particularly large. Shrinkage.
\item
  However, to fit these models, Molstad and Rothman focused on the fact that onnly a ``characteristic'' of a precision matrix is required to fit these models. ``We propose a framework to shrink a user-specified characteristic of a precision matrix estimator.''
\item
  For instance, LDA, linear regression, \ldots{}
\item
  Their approach/methodology was published in BLAH. Shrinking characteristic of precision matrix estimators.
\item
  BREAK
\item
  We wanted to explore the regression application that was mentioned but not pursued/investigated
\item
  This document will outline the research and work leading up to that
\item
  The document is outlined as follows\ldots{}
\item
  We will begin chapter 2 with a brief introduction to precision matrix estimation. This will include mentions of estimation methods/approaches and we will spend a particular amount of time elaborating on one approach which is estimation via the ADMM algorithm. This dicussion will be useful when we circle back to SCPME, the regression application, and the augmented ADMM algorithm developed in BLAH.
\item
  Lastly, the document will end with two brief tutorials for the R packages ADMMsigma and SCPME. These packages, developed by myself, will be mentioned throughout the document and were instrumental in this pursued line of research and facilitated many simulations and research direction explored throughout. Both packages have since been published on CRAN
\end{itemize}

\hypertarget{notation-and-definitions}{%
\subsection{Notation and Definitions}\label{notation-and-definitions}}

For strictly positive integers \(n\) and \(p\), we will denote \(\mathbb{R}^{n \times p}\) as the class of real matrices with dimenson \(n \times p\). The class of real, symmetric matrices with dimension \(p \times p\) will be denoted as \(\mathbb{S}^{p}\) and \(\mathbb{S}^{p}_{+}\) if we further require the object to be positive definite. The sample size in a given data set will most often correspond with \(n\) and \(p\) will correspond with the dimension of the prediction vector. If the dimension of the response vector exceeds one, we will denote it as \(r\).

Most matrices will take the form of either \(\Sigma\), which should be taken to represent the population covariance matrix, or \(\Omega\), which should be taken to represent the population \emph{precision} matrix. Note that the precision matrix is simply the inverse of the covariance matrix (\(\Omega \equiv \Sigma^{-1}\)). We will sometimes add a subscript star if the object is oracle - or known- a priori (\(\Omega_{*}\)) and the population object's estimator that minimizes/maximizes a pre-specified objective function will have a hat attached to it (\(\hat{\Omega}\)).

There will be significant matrix algebra notation throughout the document. The trace operator which sums the diagonal elements of a matrix will take the form \(tr\left(\cdot\right)\) and the \emph{exponential} trace operator will be denoted similarly as \(etr\left(\cdot\right)\). The determinant of a matrix \(\mathbf{A}\) will be denoted as \(\left|\mathbf{A}\right|\) but may also take the form as simply \(det\left(\mathbf{A}\right)\). The kronecker product of two matrices \(\mathbf{A}\) and \(\mathbf{B}\) will be denoted as \(\mathbf{A} \otimes \mathbf{B}\). Lastly, the Frobenius norm which sums the square of all entries in a matrix will be denoted as \(\left\|\mathbf{A}\right\|_{F}\) and we will define \(\left\|\mathbf{A}\right\|_{1} := \sum_{i, j}\left|\mathbf{A}_{ij}\right|\) where the \(i\)-\(j\)th element in matrix \(\mathbf{A}\) is denoted as \(\left(\mathbf{A}\right)_{ij}\) or simply \(\mathbf{A}_{ij}\).

\hypertarget{precision-matrix-estimation}{%
\chapter{Precision Matrix Estimation}\label{precision-matrix-estimation}}

\hypertarget{literature-review}{%
\subsection{Literature Review?}\label{literature-review}}

\begin{itemize}
\item
  Articles that review covariance and precision matrix estimation in high dimensional settings are \citet{pourahmadi2013high} and \citet{fan2016overview}.
\item
  The person who initially proposed the lasso-penalized gaussian likelihood precision matrix estimation is \citet{yuan2007model}.
\item
  Other authors who have proposed methods with non-lasso penalizations are (Frobenius norms) \citet{witten2009covariance}, \citet{rothman2014existence}, (non-convex penalties) \citet{lam2009sparsistency}, and \citet{fan2009network}.
\item
  Standard alternating direction method of multipliers approaches are found in \citet{boyd11distributed}.
\item
  ADMM with modification based on majorize-minimize principle is found in \citet{lange2016mm}.
\item
  Quadratic equation with ridge is solve in closed form using a similar approach to \citet{witten2009covariance}; \citet{price2015ridge}.
\item
  L1-penalized gaussian likelihood precision matrix estimator (\citet{yuan2007model}; \citet{rothman2008sparse}; \citet{friedman2008sparse}).
\end{itemize}

\hypertarget{background}{%
\section{Background}\label{background}}

Consider the case where we observe \(n\) independent, identically distributed copies of the random variable (\(X_{i}\)) where \(X_{i} \in \mathbb{R}^{p}\) is normally distributed with some mean, \(\mu\), and some variance, \(\Omega^{-1}\). That is, \(X_{i} \sim N_{p}\left( \mu, \Omega^{-1} \right)\).

Because we assume independence, we know that the probability of observing these specific observations \(X_{1}, ..., X_{n}\) is equal to

\begin{equation} 
\begin{split}
  f(X_{1}, ..., X_{n}; \mu, \Omega^{-1}) &= \prod_{i = 1}^{n}(2\pi)^{-p/2}\left| \Omega \right|^{1/2}\exp\left[ -\frac{1}{2}\left( X_{i} - \mu \right)'\Omega\left( X_{i} - \mu \right) \right] \\
  &= (2\pi)^{-np/2}\left| \Omega \right|^{n/2}\mbox{etr}\left[ -\frac{1}{2}\sum_{i = 1}^{n}\left( X_{i} - \mu \right)\left( X_{i} - \mu \right)'\Omega \right]
\end{split}
\label{eq:normallik}\notag
\end{equation}

where \(\mbox{etr}\left( \cdot \right)\) denotes the exponential trace operator. It follows that the log-likelihood for \(\mu\) and \(\Omega\) is equal to the following:

\begin{equation}
  l(\mu, \Omega | X) = const. + \frac{n}{2}\log\left| \Omega \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)'\Omega \right]
\label{eq:normalloglik}\notag
\end{equation}

If we are interested in estimating \(\mu\), it is relatively straight forward to show that the maximum likelihood estimator (MLE) for \(\mu\) is \(\hat{\mu}^{MLE} = \sum_{i = 1}^{n}X_{i}/n\) which we typically denote as \(\bar{X}\). However, in addition to \(\mu\), many applications require the estimation of \(\Sigma\) as well. We can also find a maximum likelihood estimator:

\begin{equation}
\begin{split}
\hat{\Omega}^{mle} &= \arg\max_{\Omega \in S_{+}^{p}}\left\{ const. + \frac{n}{2}\log\left| \Omega \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)'\Omega \right] \right\} \\
  &= \left[ \frac{1}{n}\sum_{i = 1}^{n}\left(X_{i} - \bar{X} \right)\left(X_{i} - \bar{X} \right)' \right]^{-1}
\end{split}
\label{eq:omegamle}\notag
\end{equation}

By setting the gradient equal to zero and plugging in the MLE for \(\mu\), we find that the MLE for \(\Omega\) is the inverse our usual sample estimator for the covariance matrix, denoted here as \(S^{-1}\). It turns out that we could have just as easily computed the maximum likelihood estimator for the precision matrix \(\Omega \equiv \Sigma^{-1}\) and taken its inverse:

\begin{equation}
\hat{\Omega}^{mle} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| \right\}
\label{eq:omegamlesimple}
\end{equation}

so that \(\hat{\Omega}^{MLE} = S^{-1}\). Beyond the formatting convenience, computing estimates for \(\Omega\) as opposed to \(\Sigma\) often poses less computational challenges -- and accordingly, the literature has placed more emphasis on efficiently solving for \(\Omega\) instead of \(\Sigma\)\footnote{Notice that here we are \emph{minimizing} the negative log-likelihood as opposed to maximizing the log-likelihood. Both procedures will result in the same estimate.}.

As in regression settings, we can construct a \emph{penalized} log-likelihood estimator by adding a penalty term, \(P\left(\Omega\right)\), to the log-likelihood so that

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + P\left( \Omega \right) \right\}
\label{eq:omegapen}
\end{equation}

\(P\left( \Omega \right)\) is often of the form \(P\left(\Omega \right) = \lambda\|\Omega \|_{F}^{2}/2\) or \(P\left(\Omega \right) = \|\Omega\|_{1}\) where \(\lambda > 0\), \(\left\|\cdot \right\|_{F}^{2}\) is the Frobenius norm and we define \(\left\|A \right\|_{1} = \sum_{i, j} \left| A_{ij} \right|\). These penalties are the ridge and lasso, respectively.

\hypertarget{admm-algorithm}{%
\section{ADMM algorithm}\label{admm-algorithm}}

Many efficient methods have been proposed to solve for \(\hat{\Omega}\) when \(\alpha = 1\). The most popular method is the graphical lasso algorithm (glasso) introduced by \citet{friedman2008sparse}. However, no methods (to the best of my knowledge) have estimated \(\Omega\) when \(\alpha \in (0, 1)\). We will use the alternating direction method of multipliers (ADMM) algorithm to do so.

As the authors state in \citet{boyd2011distributed}, the ``ADMM is an algorithm that is intended to blend the decomposability of dual ascent with the superior convergence properties of the method of multipliers.'' For our purposes, we will only focus on the ADMM algorithm but it is encouraged to read the original text from Boyd and others for a complete introduction to the other two methods.

In general, suppose we want to solve an optimization problem of the following form:

\begin{equation}
\begin{split}
  \mbox{minimize } f(x) + g(z) \\
  \mbox{subject to } Ax + Bz = c
\end{split}
\label{eq:boyd}\notag
\end{equation}

where \(x \in \mathbb{R}^{n}, z \in \mathbb{R}^{m}, A \in \mathbb{R}^{p \times n}, B \in \mathbb{R}^{p \times m}\), \(c \in \mathbb{R}^{p}\), and \(f\) and \(g\) are assumed to be convex functions (following \citet{boyd2011distributed}, the estimation procedure will be introduced in vector form though we could just as easily take \(x\) and \(z\) to be matrices). In addition to penalized precision matrix estimation, optimization problems like this arise naturally in several statistics and machine learning applications -- particularly regularization methods. For instance, we could take \(f\) to be the squared error loss, \(g\) to be the \(l_{2}\)-norm, \(c\) to be equal to zero and \(A\) and \(B\) to be identity matrices to solve the ridge regression optimization problem. In all cases, our goal is to find \(x^{*} \in \mathbb{R}^{n}\) and \(z^{*} \in \mathbb{R}^{m}\) that achieves the infimum \(p^{*}\):

\begin{equation}
p^{*} = inf\left\{ f(x) + g(z) | Ax + Bz = c \right\}
\label{eq:infimum}\notag
\end{equation}

To do so, the ADMM algorithm uses the \emph{augmented lagrangian}

\begin{equation}
L_{\rho}(x, z, y) = f(x) + g(z) + y'(Ax + Bz - c) + \frac{\rho}{2}\left\| Ax + Bz - c \right\|_{2}^{2}
\label{eq:auglagrange}\notag
\end{equation}

The algorithm consists of the following repeated iterations:

\begin{equation}
\begin{split}
  x^{k + 1} &= \arg\min_{x \in \mathbb{R}^{n}}L_{\rho}(x, z^{k}, y^{k}) \\
  z^{k + 1} &= \arg\min_{z \in \mathbb{R}^{m}}L_{\rho}(x^{k + 1}, z, y^{k}) \\
  y^{k + 1} &= y^{k} + \rho(Ax^{k + 1} + Bz^{k + 1} - c)
\end{split}  
\label{eq:admmalgo}\notag
\end{equation}

In the context of precision matrix estimation, we can let \(f\) be equal to the non-penalized likelihood, \(g\) be equal to \(P\left( \Omega \right)\), and use the constraint \(\Omega\) equal to some \(Z\) so that the lagrangian is

\begin{equation}
L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + tr\left[\Lambda\left(\Omega - Z\right)\right]
\label{eq:lagrange2}\notag
\end{equation}

and the augmented lagrangian is

\begin{equation}
L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + tr\left[\Lambda\left(\Omega - Z\right)\right] + \frac{\rho}{2}\left\|\Omega - Z\right\|_{F}^{2}
\label{eq:auglagrange2}\notag
\end{equation}

\vspace{3cm}

In the \texttt{ADMMsigma} package, we instead take

\begin{equation}
P\left( \Omega \right) = \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right]
\label{eq:penelastic}\notag
\end{equation}

so that solving the full penalized log-likelihood for \(\Omega\) results in solving

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right] \right\}
\label{eq:optimelastic}
\end{equation}

where \(0 \leq \alpha \leq 1\). This penalty, know as the \emph{elastic-net} penalty, was explored by Hui Zou and Trevor Hastie \citep{zou2005regularization} and is identical to the penalty used in the popular penalized regression package \texttt{glmnet}. Clearly, when \(\alpha = 0\) the elastic-net reduces to a ridge-type penalty and when \(\alpha = 1\) it reduces to a lasso-type penalty. Having this flexibility and generalization allows us to perform cross validation across proposed \(\alpha\) values in addition to proposed \(\lambda\) values.

The ADMM algorithm is now the following:

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \label{eq:omegaauglagrangee} \\
  Z^{k + 1} &= \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \label{eq:ZZauglagrange} \\
  \Lambda^{k + 1} &= \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right) \label{eq:lamauglagrange}
\end{align}

\hypertarget{algorithm}{%
\subsection{Algorithm}\label{algorithm}}

Set \(k = 0\) and initialize \(Z^{0}, \Lambda^{0}\), and \(\rho\). Repeat steps 1-3 until convergence:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Decompose \(S + \Lambda^{k} - \rho Z^{k} = VQV'\) via spectral decomposition\footnote{Proof of \eqref{eq:omegaalgo} in section \ref{proofomegaalgo}.}. Then
\end{enumerate}

\begin{equation}
\Omega^{k + 1} = \frac{1}{2\rho}V\left[ -Q + \left( Q^{2} + 4\rho I_{p} \right)^{1/2} \right]V'
\label{eq:omegaalgo}
\end{equation}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Elementwise soft-thresholding for all \(i = 1,..., p\) and \(j = 1,..., p\)\footnote{Proof of \eqref{eq:ZZalgo} in section \ref{proofZZalgo}.}.
\end{enumerate}

\begin{equation}
\begin{split}
Z_{ij}^{k + 1} &= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{sign}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}\right)\left( \left| \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} \right| - \lambda\alpha \right)_{+} \\
&= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{soft}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}, \lambda\alpha\right)
\label{eq:ZZalgo}
\end{split}
\end{equation}

\begin{equation}
\Lambda^{k + 1} = \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right)
\label{eq:lamalgo}\notag
\end{equation}

where \(\mbox{soft}(a, b) = \mbox{sign}(a)(\left| a \right| - b)_{+}\).

\hypertarget{scaled-form-admm}{%
\subsection{Scaled-Form ADMM}\label{scaled-form-admm}}

There is another popular, alternate form of the ADMM algorithm used by scaling the dual variable (\(\Lambda^{k}\)). Let us define \(R^{k} = \Omega - Z^{k}\) and \(U^{k} = \Lambda^{k}/\rho\).

\begin{equation}
\begin{split}
  tr\left[ \Lambda^{k}\left( \Omega - Z^{k} \right) \right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} &= tr\left[ \Lambda^{k}R^{k} \right] + \frac{\rho}{2}\left\| R^{k} \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + \Lambda^{k}/\rho \right\|_{F}^{2} - \frac{\rho}{2}\left\| \Lambda^{k}/\rho \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + U^{k} \right\|_{F}^{2} - \frac{\rho}{2}\left\| U^{k} \right\|_{F}^{2}
\label{eq:scaled}\notag
\end{split}
\end{equation}

Therefore, a scaled-form can now be written as

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega \in \mathbb{R}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + \frac{\rho}{2}\left\| \Omega - Z^{k} + U^{k} \right\|_{F}^{2} \right\} \label{eq:omegascaled}\notag \\
  Z^{k + 1} &= \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z + U^{k} \right\|_{F}^{2} \right\} \label{eq:ZZscaled}\notag \\
  U^{k + 1} &= U^{k} + \Omega^{k + 1} - Z^{k + 1} \label{eq:Uscaled}\notag
\end{align}

\begin{align}
  x^{k + 1} &= \arg\min_{x}\left\{ f(x) + \frac{\rho}{2}\left\| Ax + Bz^{k} - c + u^{k} \right\|_{2}^{2} \right\} \label{eq:xscaled}\notag \\
  z^{k + 1} &= \arg\min_{z}\left\{ g(z) + \frac{\rho}{2}\left\| Ax^{k + 1} + Bz - c + u^{k} \right\|_{2}^{2} \right\} \label{eq:zscaled}\notag \\
  u^{k + 1} &= u^{k} + Ax^{k + 1} + Bz^{k + 1} - c \label{eq:uscaled}\notag
\end{align}

Note that there are limitations to using this method. Because the dual variable is scaled by \(\rho\) (the step size), this form limits one to using a constant step size without making further adjustments to \(U^{k}\). It has been shown in the literature that a dynamic step size (like the one used in \texttt{ADMMsigma}) can significantly reduce the number of iterations required for convergence.

\hypertarget{stopping-criterion}{%
\subsection{Stopping Criterion}\label{stopping-criterion}}

In discussing the optimality conditions and stopping criterion, we will follow the steps outlined in \citet{boyd2011distributed} and cater them to precision matrix estimation.

Below we have three optimality conditions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Primal:
\end{enumerate}

\begin{equation}
\Omega^{k + 1} - Z^{k + 1} = 0
\label{eq:admmprimal}\notag
\end{equation}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Dual:
\end{enumerate}

\begin{equation}
\begin{split}
  0 &\in \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k + 1} \\
  0 &\in \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1}
\label{eq:admmdual}\notag
\end{split}
\end{equation}

The first dual optimality condition is a result of taking the sub-differential of the lagrangian (non-augmented) with respect to \(\Omega^{k + 1}\) (note that we must honor the symmetric constraint of \(\Omega^{k + 1}\)) and the second is a result of taking the sub-differential of the lagrangian with respect to \(Z^{k + 1}\) (no symmetric constraint).

We will define the left-hand side of the first condition as the primal residual \(r^{k + 1} = \Omega^{k + 1} - Z^{k + 1}\). At convergence, optimality conditions require that \(r^{k + 1} \approx 0\). The second residual we will define is the dual residual \(s^{k + 1} = \rho\left( Z^{k + 1} - Z^{k} \right)\). This residual is derived from the following:

Because \(\Omega^{k + 1}\) is the minimizer of the augmented lagragian\footnote{Proof of \eqref{eq:dualresidual} in section \ref{proofdualresidual}.}, this implies that \(0 \in \partial L_{p}\left( \Omega, Z^{k}, \Lambda^{k} \right)\) which further implies that

\begin{equation}
0 \in \rho\left( Z^{k + 1} - Z^{k} \right)
\label{eq:dualresidual}
\end{equation}

Like the primal residual, at convergence optimality conditions require that \(s^{k + 1} \approx 0\). Note that the second dual optimality condition is always satisfied\footnote{Proof of \eqref{eq:dualopt} in section \ref{proofdualopt}.}:

\begin{equation}
\begin{split}
  0 \in \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1} \\
\label{eq:dualopt}
\end{split}
\end{equation}

One possible stopping criterion is to set \(\epsilon^{rel} = \epsilon^{abs} = 10^{-3}\) and stop the algorithm when \(\epsilon^{pri} \leq \left\| r^{k + 1} \right\|_{F}\) and \(\epsilon^{dual} \leq \left\| s^{k + 1} \right\|_{F}\) where

\begin{equation}
\begin{split}
  \epsilon^{pri} &= p\epsilon^{abs} + \epsilon^{rel}\max\left\{ \left\| \Omega^{k + 1} \right\|_{F}, \left\| Z^{k + 1} \right\|_{F} \right\} \\
  \epsilon^{dual} &= p\epsilon^{abs} + \epsilon^{rel}\left\| \Lambda^{k + 1} \right\|_{F}
\label{eq:admmstopping}\notag
\end{split}
\end{equation}

\hypertarget{simulations}{%
\section{Simulations}\label{simulations}}

In the simulations below, we generate data from a number of different oracle precision matrices with various structures. For each data-generating procedure, the \texttt{ADMMsigma()} function was run using 5-fold cross validation. After 20 replications, the cross validation errors were totalled and the optimal tuning parameters were selected (results below the top figure). These results are compared with the Kullback Leibler (KL) losses between the estimates and the oracle precision matrix (bottom figure).

We can see below that our cross validation procedure consistently chooses tuning parameters that are close to the optimal parameters with respsect to the oracle.

HOW DID WE GENERATE THE DATA?

\hypertarget{plots}{%
\subsection{Plots}\label{plots}}

\hypertarget{compound-symmetric-p-100-n-50}{%
\subsubsection{Compound Symmetric: P = 100, N = 50}\label{compound-symmetric-p-100-n-50}}

@@

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth,]{images/compound_N50_P100}

\caption{Figure caption here.}\label{fig:admmsim1}
\end{figure}

\hypertarget{compound-symmetric-p-10-n-1000}{%
\subsubsection{Compound Symmetric: P = 10, N = 1000}\label{compound-symmetric-p-10-n-1000}}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth,]{images/compound_N1000_P10}

\caption{Figure caption here.}\label{fig:admmsim2}
\end{figure}

--\textgreater{}
 --\textgreater{}
 --\textgreater{}
 --\textgreater{}
 --\textgreater{}

--\textgreater{}
 --\textgreater{}
 --\textgreater{}
 --\textgreater{}

\hypertarget{tridiagonal-p-100-n-50}{%
\subsubsection{Tridiagonal: P = 100, N = 50}\label{tridiagonal-p-100-n-50}}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth,]{images/repsKLtridiag_N50_P100}

\caption{Figure caption here.}\label{fig:admmsim5}
\end{figure}

\hypertarget{scpme}{%
\chapter{SCPME}\label{scpme}}

\hypertarget{literature-review-1}{%
\subsection{literature review?}\label{literature-review-1}}

\begin{itemize}
\item
  A similar estimator was proposed by \citet{dalal2017sparse} when \(C = 0\).
\item
  \begin{itemize}
  \tightlist
  \item
    Quadratic equation with ridge is solve in closed form using a similar approach to \citet{witten2009covariance}; \citet{price2015ridge}.
  \end{itemize}
\end{itemize}

\hypertarget{other-stuff}{%
\subsection{other stuff}\label{other-stuff}}

The penalty proposed by \citet{molstad2017shrinking} is one of the following form:

\begin{equation}
P\left(\Omega\right) = \lambda\left\| A\Omega B - C \right\|_{1}
\label{eq:pen2}
\end{equation}

where \(A \in \mathbb{R}^{m \times p}, B \in \mathbb{R}^{p \times q}, \mbox{ and } C \in \mathbb{R}^{m \times q}\) are matrices assumed to be known and specified by the user. Solving the full penalized log-likelihood for \(\Omega\) results in solving

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + \lambda\left\| A\Omega B - C \right\|_{1} \right\}
\label{eq:omegaloglik2}
\end{equation}

This form of penalty is particularly useful because matrices \(A, B, \mbox{ and } C\) can be constructed so that we penalize the sum, absolute value of a \emph{characteristic} of the precision matrix \(\Omega\). This type of penalty leads to many new, interesting, and novel estimators for \(\Omega\). An example of one such estimator (suppose we observe \(n\) samples of \(Y_{i} \in \mathbb{R}^{r}\)) would be one where we set \(A = I_{p}, B = \Sigma_{xy}, \mbox{ and } C = 0\) where \(\Sigma_{xy}\) is the covariance matrix of \(X\) and \(Y\). This penalty has the effect of assuming sparsity in the forward regression coefficient \(\beta \equiv \Omega\Sigma_{xy}\). Of course, in practice we do not know the true covariance matrix \(\Sigma_{xy}\) but we might consider using the sample estimate \(\hat{\Sigma}_{xy} = \sum_{i = 1}^{n}\left(X_{i} - \bar{X}\right)\left(Y_{i} - \bar{Y}\right)^{T}/n\)

We will explore how to solve for \(\hat{\Omega}\) in the next section.

\hypertarget{augmented-admm-algorithm}{%
\section{Augmented ADMM Algorithm}\label{augmented-admm-algorithm}}

\emph{This section requires general knowledge of the alternating direction method of multipliers (ADMM) algorithm. I would recommend reading this overview I have written \href{https://mgallow.github.io/ADMMsigma/articles/Details.html\#admm-algorithm}{here} before proceeding.}

The ADMM algorithm - thanks to it's flexibility - is particularly well-suited to solve penalized-likelihood optimization problems that arise naturally in several statistics and machine learning applications. Within the context of \citet{molstad2017shrinking}, this algorithm would consist of iterating over the following three steps:

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}L_{\rho}(\Omega, Z^{k}, \Lambda^{k}) \label{eq:omegaauglagrange2}\notag \\
  Z^{k + 1} &= \arg\min_{Z \in \mathbb{R}^{n \times r}}L_{\rho}(\Omega^{k + 1}, Z, \Lambda^{k}) \label{eq:ZZauglagrange2}\notag \\
  \Lambda^{k + 1} &= \Lambda^{k} + \rho\left(A\Omega^{k + 1}B - Z^{k + 1} - C \right) \label{eq:lamauglagrange2}\notag
\end{align}
